---
title: "Polynomial fitting"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(ggplot2)
library(dplyr)
```


## Generate some data from a known 3rd-degree polynomial function

Using the equation:
y = ax^3 + bx^2 + cx + d

Set a and d to arbitrary positive numbers and b, c to 0.
Choose an x range from 0 to 100 in 0.1 increments.
Calculate and plot y.

```{r arbitrary_equation}
a <- 2
b <- 0
c <- 0
d <- 6
x <- seq(0,100,0.1)
y <- a*x^3 + b*x^2 + c*x + d
df_arbitrary <- bind_cols(x=x, y=y)
sample_plot <- ggplot(df_arbitrary, aes(x=x, y=y)) + geom_point(alpha=0.3, color = "blue") + ggtitle("Arbitrary 3rd-degree polynomial")
sample_plot
```

## Example of fitting polynomials from R docs

The lm function can be used to fit polynomials of varying degree.

Here 4 different best fits are determined for the relationship between stopping distance and speed using the built-in cars dataset.

ANOVA is used to evaluate the best fitting function.


```{r cars}
## An example of polynomial regression
plot(cars, xlab = "Speed (mph)", ylab = "Stopping distance (ft)",
    las = 1, xlim = c(0, 25))

# generate x values for fit lines
d_speed <- seq(0, 25, length.out = 200) 

# fit different degree polynomials
for(degree in 1:4) {
  fm <- lm(dist ~ poly(speed, degree, raw=TRUE), data = cars)
  # create variable to hold results
  assign(paste("cars", degree, sep = "."), fm) 
  lines(d_speed, predict(fm, data.frame(speed = d_speed)), col = degree)
}
anova(cars.1, cars.2, cars.3, cars.4)
```

## Applying polynomial fitting to arbitrary data

### 1. Leaving coefficents unspecified

Here we use poly(x, 3, raw=TRUE) to restrict the fit to a 3rd-degree polynomial and recover the "raw" coefficients (otherwise we would get coefficients for the "orthogonal polynomial" which we do not want here)

The summary shows that the middle coefficients (b and c) are correctly determined to be close to 0.

```{r lm_arbit}

## Try to fit without specifying coefficients

plot(df_arbitrary, xlab = "x", ylab = "y",
    las = 1)

mypoly <- poly(x, 3, raw = TRUE)
fit_open <- lm(y ~ mypoly, data = df_arbitrary)

# add the fit line to the plot and print the summary
lines(x, predict(fit_open, data.frame(x = x)), col = "red")
summary(fit_open)
```

### 2. Constraining the middle coefficients to be 0

Using the "I" notation we can specify the model to be of the form ax^3 + d. (The intercept is implied in the I notation; if we wanted to constrain that to be 0 as well, we could say lm(y ~ I(x^3) ~ 1))


```{r lm_arbit_specify}
plot(df_arbitrary, xlab = "x", ylab = "y",
    las = 1)
fit_spec <- lm(y ~ I(x^3), data = df_arbitrary)
lines(x, predict(fit_spec, data.frame(x = x)), col = "yellow")
summary(fit_spec)
```

### Trying it on more realistic data

The dataset size is reduced to 100 and gaussian noise with a SD of 100 is added

The fit is still determined with good confidence in the coefficient, but less confidence in the intercept (perhaps because noise disproportionately affects smaller values of y)

```{r lm_arbit_noisy}

# Add noise and reduce size of dataset
noise <- rnorm(100, mean=0, sd = 100)
noisy_y <- df_arbitrary$y[1:100] + noise

df_noisy <- bind_cols(x = df_arbitrary$x[1:100], y = noisy_y)
plot(df_noisy, xlab = "x", ylab = "y",
    las = 1)
fit_noisy <- lm(y ~ I(x^3), data = df_noisy)
lines(x, predict(fit_noisy, data.frame(x = x)), col = "orange")
summary(fit_noisy)
coefficients(fit_noisy)
```


Load data from inclusions.
The column names are the cytoplasmic mean intensity.

```{r load inclusion data}
library(readxl)
IB_volumes_for_R <- read_excel("~/Desktop/research/images/images IB growth/htt tiffs 18_3_6/Done/IB volumes for R.xlsx", 
    na = "NA")
```

Fit data from a single inclusion at a time.

```{r fit single column}
fit_incl1 <- lm(`1392` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`2695` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`1857` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`1686` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`1616` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`2130` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`2010` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`1981` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`2654` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`1272` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

fit_incl1 <- lm(`1782.3744799999999` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)


# plot one inclusion's data
fit_incl1 <- lm(`1392` ~ I(time^3), data = IB_volumes_for_R)
summary(fit_incl1)

incl_1392 <- ggplot(IB_volumes_for_R, aes(x=time, y=`1392`)) + geom_point(alpha=0.3, color = "blue") + ggtitle("Inclusion 1392")

incl_1392
```

Fit data from the whole table and save the coefficients, R^2, and p values in a new table.

``` {r apply fit to several columns}

# the [,2:ncol()] means process all rows, and process column 2 (the first inclusion data) to the end
# the "2" argument means apply to each column

fit_summs <- apply(IB_volumes_for_R[,2:ncol(IB_volumes_for_R)], 2, function(x) summary(lm(x ~ I(time^3), data = IB_volumes_for_R)))

# to get the values in a usable form
# fit_summs$`1392`["coefficients"] gives a table where the rows are the coeffs, 1st col is the coeff itself, 4th col is the p value

# we want the column name, R squared, coefficients (intercept, I(time^3)), and for each coefficient the P value.

names(fit_summs) <- colnames(IB_volumes_for_R)[2:ncol(IB_volumes_for_R)]

fit_results <- lapply(seq_along(fit_summs), function(x) {
  data.frame(cyto <- names(fit_summs)[x],
  Intercept <- fit_summs[[x]]$coefficients[1,1],
  A <- fit_summs[[x]]$coefficients[2,1],
  Int_P <- fit_summs[[x]]$coefficients[1,4],
  A_P <- fit_summs[[x]]$coefficients[2,4],
  rSquared <- fit_summs[[x]]$r.squared,
  adjRSquared <- fit_summs[[x]]$adj.r.squared)
})

fits <- do.call(rbind, fit_results)
write.csv(fits, file =  paste("~/Desktop/fits", Sys.time(),".csv"))

```