---
title: "Summarizing Overlap Volume from Imaris Statistics"
output: html_notebook
---

# Overview
This notebook calculates total volume overlap of two Surfaces defined in Imaris. It also calculates each surface's total volume and normalizes the volume overlap to each total.

# Experiment details
ENTER YOUR EXPERMIMENT DETAILS HERE
Cell type, conditions, stains...

# Analysis details
ENTER YOUR SURFACE INFO HERE
Surfaces 1 is ----.
Surfaces 2 is ----.

```{r setup}
knitr::opts_chunk$set(
	message = FALSE,
	warning = TRUE,
	include = FALSE
)
# ---- Setup ----
require(tidyverse) # organizing data
require(readr) # reading data
require(stringr) # processing strings
require(ggplot2)  # plotting
require(rstudioapi) # selecting output
```

# Input data
The input files are Imaris single or batch output tables. 

*You will need two Volume tables and one Overlapped Volume to Surfaces table.*

Each table is a CSV file containing one type of measurement (such as volume) for many objects within one or many images. 

Each observation (row) represents one object detected by Imaris. Some may be clumps.  

The channel is read from the Original Component Name column.

# Loading data files

##TODO: Open a folder of files matching a pattern

First, load the Surfaces 1 (S1) Volume data.

```{r load S1 volume, echo=FALSE}
# ---- Prompt for an object volume file ----

S1VolFile <- file.choose()
print("Surfaces 1 volume file:")
S1VolFile # report the file that was selected

# Read the data from the file
S1VolData <- read_csv(S1VolFile,
                       col_types = cols(...12 = col_skip()), 
                       skip = 3,
                       locale = locale())

```

Next, load the Surfaces 2 (S2) Volume data.

```{r load S2 volume, echo=FALSE}
# ---- Prompt for an object volume file ----

S2VolFile <- file.choose()
print("Surfaces 2 volume file:")
S2VolFile # report the file that was selected

S2VolData <- read_csv(S2VolFile,
                       col_types = cols(...12 = col_skip()), 
                       skip = 3,
                       locale = locale())

```
# Finally, load one of the Overlap Volume files.
```{r load S2 in S1 overlap data}

overlapFile <- file.choose()
overlapFile

overlapMLData <- read_csv(overlapMLFile1,
                       col_types = cols(...13 = col_skip()), 
                       skip = 3,
                       locale = locale())
```

# Calculation of total S1 and S2 volume per image  

The Volume data is summarized by file name.

```{r calculate total S1 and S2 volume per image}

S1VolSums <- S1VolData %>% 
  group_by(`Original Image Name`,`Original Image ID`) %>% 
  summarise(nS1Objects = n(),
            SumS1Volume = sum(Volume))

S2VolSums <- S2VolData %>% 
  group_by(`Original Image Name`, `Original Image ID`) %>% 
  summarise(nS2Objects = n(),
            SumS2Volume = sum(Volume))

```

# Calculation of total overlap (S2 in S1) volume per image  

The overlap volume is stored in the Overlapped Volume to Surfaces column. There are 2 tables that contain this data, one from each Surface, but the numbers are identical because they are simply the volume of the intersection of the two Surfaces.  

```{r sum overlap data}

overlapMLFile1 <- file.choose()
overlapMLFile1

overlapMLData <- read_csv(overlapMLFile1,
                       col_types = cols(...13 = col_skip()), 
                       skip = 3,
                       locale = locale())
# 
# overlapMLFile2 <- file.choose()
# overlapMLFile2
# 
# overlapMLData2 <- read_csv(overlapMLFile2,
#                        col_types = cols(...13 = col_skip()), 
#                        skip = 3,
#                        locale = locale())
# 
# overlapMLData <- rbind(overlapMLData1, overlapMLData2)

overlapMLSums <- overlapMLData %>% 
  group_by(`Original Image Name`,`Original Image ID`) %>% 
  summarise(SumS2InS1 = sum(`Overlapped Volume to Surfaces`))

overlapMLSums

```


# Calculation of total overlap (S1 in S2) volume per image  

The amount of S1 volume that is in S2 is tabulated in  "Surfaces_1_Overlapped_Volume_to_Surfaces_Surfaces=Tom20-405". This data is loaded and summarized by file name.

```{r load S1 in S2 overlap data}

overlapLMFile1 <- file.choose()
overlapLMFile1

overlapLMData <- read_csv(overlapLMFile1,
                       col_types = cols(...13 = col_skip()), 
                       skip = 3,
                       locale = locale())

# overlapLMFile2 <- file.choose()
# overlapLMFile2
# 
# overlapLMData2 <- read_csv(overlapLMFile2,
#                        col_types = cols(...13 = col_skip()), 
#                        skip = 3,
#                        locale = locale())
# 
# overlapLMData <- rbind(overlapLMData1, overlapLMData2)

overlapLMSums <- overlapLMData %>% 
  group_by(`Original Image Name`,`Original Image ID`) %>% 
  summarise(SumS1InS2 = sum(`Overlapped Volume to Surfaces`))

overlapLMSums

```
# Calculation of fractional overlap per image

The volume and overlap tables are merged by file name.

The date and slide number are extracted from the original file name and path, and the data are summarized by date and blinded slide number, which can be cross-referenced to the unblinded slide information.

```{r merge tables}

mergedVolTable <- inner_join(S1VolSums, S2VolSums, by=NULL)
mergedTable <- inner_join(mergedVolTable, overlapMLSums, by = NULL)
mergedTable <- inner_join(mergedTable, overlapLMSums, by = NULL)

mergedTable <- mergedTable %>%
  mutate(`Overlap/S2Vol` = SumS2InS1/SumS2Volume,
         `Overlap/S1Vol` = SumS1InS2/SumS1Volume,
         slideNum = str_sub(`Original Image Name`, 1, 1),
         imageDate = str_extract(`Original Image ID`, "2023-[0-9]{2}-[0-9]{2}"))

mergedTable
```

Then the fractional overlap is calculated on a per-image basis (this is different from the Overlap Ratio table produced by Imaris, which is on a per-object basis)


```{r summarise}
summaryTable <- mergedTable %>% 
  group_by(imageDate, slideNum) %>%
  summarise(nImages = n(),
            TotalS2 = sum(SumS2Volume),
            TotalS1 = sum(SumS1Volume),
            TotalS2OverlapWithS1 = sum(SumS2InS1),
            TotalS1OverlapWithS2 = sum(SumS1InS2)) %>%
  mutate(`WtdAvg_Overlap/S2Vol` = TotalS2OverlapWithS1/TotalS2,
         `WtdAvg_Overlap/S1Vol` = TotalS1OverlapWithS2/TotalS1)

summaryTable
```

# Plotting results

Box plots are created for organelle volume and overlap volume.
Each image is one data point.
All points are plotted.

## Box-and-whisker plot design
- A small white square is plotted at the mean.
- Lower and upper hinges correspond to the first and third quartiles (the 25th and 75th percentiles). 
- The whiskers extends from the hinge to the most extreme value no further than 1.5 times the IQR from the hinge.
-- IQR is the inter-quartile range, or distance between the first and third quartiles.

```{r plot overlaps}

# plot overlap fraction of S2 that is overlapped with S1

box_overlapML <- ggplot(mergedTable, 
                      aes(x=slideNum, y=`Overlap/S2Vol`)) + 
  geom_boxplot() +
  geom_dotplot(binaxis='y', 
               stackdir='center', 
               dotsize = .5, 
               fill="red") +
  stat_summary(fun=mean, geom="point", shape=1, size=3) +
  labs(title = "S2phagy", subtitle = paste("Images collected starting",mergedTable$imageDate[1])) +
  xlab("Slide number") +
  ylab("Overlap volume/S2 volume")

box_overlapML

# plot overlap fraction of S1 that is overlapped with S2

box_overlapLM <- ggplot(mergedTable, 
                      aes(x=slideNum, y=`Overlap/S1Vol`)) + 
  geom_boxplot() +
  geom_dotplot(binaxis='y', 
               stackdir='center', 
               dotsize = .5, 
               fill="red") +
  stat_summary(fun=mean, geom="point", shape=1, size=3) +
  labs(title = "S2phagy", subtitle = paste("Images collected starting",mergedTable$imageDate[1])) +
  xlab("Slide number") +
  ylab("Overlap volume/S1 volume")

box_overlapLM
```

```{r plot volumes}
box_S2 <- ggplot(mergedTable, 
                      aes(x=slideNum, y=SumS2Volume)) + 
  geom_boxplot() +
  geom_dotplot(binaxis='y', 
               stackdir='center', 
               dotsize = .5, 
               fill="red") +
  stat_summary(fun=mean, geom="point", shape=1, size=3) +
  labs(title = "S2 volume", subtitle = paste("Images collected starting",mergedTable$imageDate[1])) +
  xlab("Slide number") +
  ylab("S2 volume per image")

box_S2


box_S1 <- ggplot(mergedTable, 
                      aes(x=slideNum, y=SumS1Volume)) + 
  geom_boxplot() +
  geom_dotplot(binaxis='y', 
               stackdir='center', 
               dotsize = .5, 
               fill="red") +
  stat_summary(fun=mean, geom="point", shape=1, size=3) +
  labs(title = "S1 volume", subtitle = paste("Images collected starting",mergedTable$imageDate[1])) +
  xlab("Slide number") +
  ylab("S1 volume per image")

box_S1

```

# Output

The merged table and summary table are saved as CSV files.
Plots are saved as PDF.

Select an output directory when running this code.

```{r save files}

outputDir <- rstudioapi::selectDirectory(caption = "Select the output directory", label = "Select Folder")
outputTime <- format(Sys.time(), "%m%e%Y-%H%M%S")
outputMergedFile = paste(outputTime,"volume merged.csv")
outputSummaryFile = paste(outputTime,"volume summary.csv")

write_csv(mergedTable,file.path(outputDir, outputMergedFile))
write_csv(summaryTable,file.path(outputDir, outputSummaryFile))

ggsave(filename = paste(outputTime, "overlap over S2.pdf"), plot = box_overlapML, path = outputDir)
ggsave(filename = paste(outputTime, "overlap over S1.pdf"), plot = box_overlapLM, path = outputDir)

ggsave(filename = paste(outputTime, "S2volImage.pdf"), plot = box_S2, path = outputDir)

ggsave(filename = paste(outputTime, "S1volImage.pdf"), plot = box_S1, path = outputDir)

```

