---
title: "Imaging Time Lags with Real STORM Data"
author: "Theresa Swayne"
date: "3/3/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(readxl)
library(readr)
library(dplyr)
```

## STORM data

Data was taken from Alex's datasets on the computer from November 2019 to February 2020.

For most experiments full metadata was saved and the "Recorded Data" sheet exported as CSV.

The time lag between successive frames was calculated from the metadata.

```{r load_STORM}
mydir <- "~/Documents/home-github/data-analysis/R/Green scope timing/2020-03-01 green timing/Alex metadata/" 
storm <- read_csv(file.path(mydir,"2019-11-14 231_1hr_spread_1to50(12g10)_1to200(mab13)_13.csv"))
storm <- storm[-c(1),c(1:11)] # remove blank row
storm$TimeLag <- c(NA, diff(storm$`Time [s]`)) # calculate time difference from one row to the next
storm <- na.omit(storm) # remove first row that contains an NA for the diff
```

Summary statistics:

```{r stats}
summarise(storm,mean=mean(TimeLag), stdev = sd(TimeLag), median = median(TimeLag), min=min(TimeLag), max=max(TimeLag))
```

Histogram:

```{r plotlags1, echo=FALSE}
#histo <- ggplot(storm, aes(x = TimeLag)) + geom_histogram(bins = 50) + ggtitle("STORM time intervals")

loghisto<- ggplot(storm, aes(x = TimeLag)) + geom_histogram() + scale_y_log10() + ggtitle("STORM time intervals, log scale")

#histo
loghisto

```

Unlike in the earlier test data, in this real experiment the longer time lags appear in a different pattern. 

1. The vast majority of time lags are ~0.05 s
2. A few hundred are about twice that, ~0.08 s
3. A small number are irregular intervals, a very few much longer, and concentrated towards the last third of the experiment.

```{r plotlags2, echo=FALSE}
lag_vs_index <- ggplot(storm, aes(x=Index, y=TimeLag)) + geom_point(alpha=0.3, color = "blue") + ggtitle("Time lag vs. timepoint")
lag_vs_index
```

However, this does not add much to the length of the experiment in this case.

```{r expectedtime min}
expected <- median(storm$TimeLag) * length(storm$TimeLag) /  60
paste("Median time * # timepoints =",expected, "s expected time required")
actual <- sum(storm$TimeLag)/60
paste("Sum of time intervals =", actual, "min actual time taken")
delay <- actual - expected
paste("Actual time - expected time =",delay,"min added to experiment")
```

It seems like the longer time lags come up at regular intervals at the beginning of the experiment.

```{r plotlags3, echo=FALSE}
lag_vs_index_subset <- ggplot(storm[1:1000,], aes(x=Index, y=TimeLag)) + geom_point(alpha=0.3, color = "blue") + ggtitle("Time lag vs. timepoint, 1st 1000 frames")
lag_vs_index_subset
```

And although the time is not that constant at the end of the experiment, the interval seems fairly so.

```{r plotlags4, echo=FALSE}
lag_vs_index_subset2 <- ggplot(storm[28999:29999,], aes(x=Index, y=TimeLag)) + geom_point(alpha=0.3, color = "blue") + ggtitle("Time lag vs. timepoint, last 1000 frames")
lag_vs_index_subset2
```

TODO: 
Extract the rows with lag between 0.05 and 0.10.
Sort by index and then add a new column containing the diff between indices.
Get summary stats and histo for this new diff column.
If the distribution is tight then there is a periodic increase in the time required which must be due to some hw/sw event.

