---
title: "Imaging Time Lags"
author: "Theresa Swayne"
date: "2/28/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(readxl)
library(dplyr)
```

## 3-channel STORM experiment

A 3-channel STORM expt was run with exposure time "1 frame" for 30000 frames.
The time lag between successive frames was calculated from the metadata.

```{r load_3chSTORM}
mydir <- "~/Desktop/" 
storm3ch <- read_excel(file.path(mydir,"2020-02-29 green timing/storm001.xlsx"))
storm3ch <- storm3ch[-c(1),c(1:11)] # remove blank row
storm3ch$TimeLag <- c(NA, diff(storm3ch$`Time [s]`)) # calculate time difference from one row to the next
```

The values have a wide range.

```{r stats}
summarise(storm3ch,mean=mean(TimeLag, na.rm = TRUE), stdev = sd(TimeLag, na.rm=TRUE), median = median(TimeLag, na.rm = TRUE), min=min(TimeLag,na.rm = TRUE), max=max(TimeLag,na.rm = TRUE))
```

Most time intervals are short:

```{r plotlags, echo=FALSE}
histo <- ggplot(storm3ch, aes(x = TimeLag)) + geom_histogram(bins = 50, na.rm=TRUE) + ggtitle("3-channel STORM time intervals")
histo
```

But there are a large number of timepoints where there is a longer lag between points:

```{r plotlags2, echo = FALSE}
loghisto<- ggplot(storm3ch, aes(x = TimeLag)) + geom_histogram() + scale_y_log10() + ggtitle("3-channel STORM time intervals, log scale")
loghisto
```

And these occur throughout the experiment:
```{r plotlags3ch, echo=FALSE}
lag_vs_index <- ggplot(storm3ch, aes(x=Index, y=TimeLag)) + geom_point(alpha=0.3, color = "blue") + ggtitle("3-channel STORM lag vs. timepoint")
lag_vs_index
```

However, this lag added only slightly to the length of the experiment:
```{r expectedtime}
expected <- median(storm3ch$TimeLag, na.rm=TRUE) * length(storm3ch$TimeLag)
paste("Median time * # timepoints =",expected, "s expected time required")
actual <- sum(storm3ch$TimeLag, na.rm=TRUE)
paste("Sum of time intervals =", actual, "s actual time taken")
delay <- (actual - expected)/60
paste("Actual time - expected time =",delay,"min added in 30-min experiment")
```

## 1-channel STORM experiment

A 1-channel (647 nm) STORM experiment was run with all other  parameters the same.
```{r load_1chSTORM}
storm1ch <- read_excel(file.path(mydir,"2020-02-29 green timing/storm 647 only.xlsx"))
storm1ch <- storm1ch[-c(1),c(1:11)] # remove blank row
storm1ch$TimeLag <- c(NA, diff(storm1ch$`Time [s]`)) # calculate time difference from one row to the next
```

Based on the histogram and the distribution throughout the experiment, the elimination of channels does not affect the lagging behavior.
```{r plotlags1ch, echo = FALSE}
loghisto<- ggplot(storm1ch, aes(x = TimeLag)) + geom_histogram() + scale_y_log10() + ggtitle("1-channel STORM time intervals, log scale")
lag_vs_index <- ggplot(storm1ch, aes(x=Index, y=TimeLag)) + geom_point(alpha=0.3, color = "blue") + ggtitle("1-channel STORM lag vs. timepoint")
loghisto
lag_vs_index
```

Although the mean interval is lower as expected, the standard deviation is higher.

The total delay is twice as long in the 1-channel experiment, but still only ~3% of the total:

```{r stats1ch}
summarise(storm1ch,mean=mean(TimeLag, na.rm = TRUE), stdev = sd(TimeLag, na.rm=TRUE), median = median(TimeLag, na.rm = TRUE), min=min(TimeLag,na.rm = TRUE), max=max(TimeLag,na.rm = TRUE))
expected1ch <- median(storm1ch$TimeLag, na.rm=TRUE) * length(storm1ch$TimeLag)
paste("Median time * # timepoints =",expected1ch, "s expected time required")
actual1ch <- sum(storm1ch$TimeLag, na.rm=TRUE)
paste("Sum of time intervals =", actual1ch, "s actual time taken")
delay1ch <- (actual1ch - expected1ch)/60
paste("Actual time - expected time =",delay1ch,"min added in 30-min 1-channel experiment")
```


## Andor EMCCD without STORM

A triggered TIRF experiment was run with all imaging parameters the same but only 1000 frames.

```{r load_tirf}
trigtirf <- read_excel(file.path(mydir,"2020-02-29 green timing/triggered tirf.xlsx"))
trigtirf  <- trigtirf[-c(1),c(1:11)] # remove blank row
trigtirf$TimeLag <- c(NA, diff(trigtirf$`Time [s]`)) # calculate time difference from one row to the next
```

Here, the time intervals were much more consistent.

```{r plotlags_trigtirf, echo = FALSE}
loghisto<- ggplot(trigtirf, aes(x = TimeLag)) + geom_histogram() + scale_y_log10() + ggtitle("Triggered TIRF time intervals, log scale")
lag_vs_index <- ggplot(trigtirf, aes(x=Index, y=TimeLag)) + geom_point(alpha=0.3, color = "blue") + ggtitle("Triggered TIRF lag vs. timepoint")
loghisto
lag_vs_index
```

In a straight TIRF experiment, as expected from the graphs, the delay is negligible.


```{r stats_trigtirf}
summarise(trigtirf,mean=mean(TimeLag, na.rm = TRUE), stdev = sd(TimeLag, na.rm=TRUE), median = median(TimeLag, na.rm = TRUE), min=min(TimeLag,na.rm = TRUE), max=max(TimeLag,na.rm = TRUE))
expected_trigtirf <- median(trigtirf$TimeLag, na.rm=TRUE) * length(trigtirf$TimeLag)
paste("Median time * # timepoints =",expected_trigtirf, "s expected time required")
actual_trigtirf <- sum(trigtirf$TimeLag, na.rm=TRUE)
paste("Sum of time intervals =", actual_trigtirf, "s actual time taken")
delay_trigtirf <- (actual_trigtirf - expected_trigtirf)/60
paste("Actual time - expected time =",delay_trigtirf,"min added in 1000-frame experiment")
```


## 4-channel triggered spinning-disk experiments


A triggered spinning-disk experiment was run with 1-frame exposure and 4 channels, 680 frames.

```{r load_sd1, echo=FALSE}
sd1 <- read_excel(file.path(mydir,"2020-02-29 green timing/4ch triggered 001.xlsx"))
sd1  <- sd1[-c(1),c(1:11)] # remove blank row
sd1$TimeLag <- c(NA, diff(sd1$`Time [s]`)) # calculate time difference from one row to the next
```

Here we also had extremely low variability.

```{r plotlags_sd1, echo = FALSE}
loghisto<- ggplot(sd1, aes(x = TimeLag)) + geom_histogram() + scale_y_log10() + ggtitle("Triggered SD time intervals, log scale")
lag_vs_index <- ggplot(sd1, aes(x=Index, y=TimeLag)) + geom_point(alpha=0.3, color = "blue") + ggtitle("Triggered SD lag vs. timepoint")
loghisto
lag_vs_index
```


```{r stats_sd1, echo=FALSE}
summarise(sd1,mean=mean(TimeLag, na.rm = TRUE), stdev = sd(TimeLag, na.rm=TRUE), median = median(TimeLag, na.rm = TRUE), min=min(TimeLag,na.rm = TRUE), max=max(TimeLag,na.rm = TRUE))
expected_sd1 <- median(sd1$TimeLag, na.rm=TRUE) * length(sd1$TimeLag)
paste("Median time * # timepoints =",expected_sd1, "s expected time required")
actual_sd1 <- sum(sd1$TimeLag, na.rm=TRUE)
paste("Sum of time intervals =", actual_sd1, "s actual time taken")
delay_sd1 <- (actual_sd1 - expected_sd1)/60
paste("Actual time - expected time =",delay_sd1,"min added in 680-frame experiment")
```


Another triggered spinning-disk experiment was run with 1-frame exposure and 4 channels, with 128x128 ROI, for 2000 frames. Results are similar.

```{r load_sd128, echo=FALSE}
sd128 <- read_excel(file.path(mydir,"2020-02-29 green timing/4ch triggered 128.xlsx"))
sd128  <- sd128[-c(1),c(1:11)] # remove blank row
sd128$TimeLag <- c(NA, diff(sd128$`Time [s]`)) # calculate time difference from one row to the next
```

```{r plotlags_sd128, echo = FALSE}
loghisto<- ggplot(sd128, aes(x = TimeLag)) + geom_histogram() + scale_y_log10() + ggtitle("Triggered SD 128x128 time intervals, log scale")
lag_vs_index <- ggplot(sd128, aes(x=Index, y=TimeLag)) + geom_point(alpha=0.3, color = "blue") + ggtitle("Triggered SD 128 x 128 lag vs. timepoint")
loghisto
lag_vs_index
```

```{r stats_sd128, echo=FALSE}
summarise(sd128,mean=mean(TimeLag, na.rm = TRUE), stdev = sd(TimeLag, na.rm=TRUE), median = median(TimeLag, na.rm = TRUE), min=min(TimeLag,na.rm = TRUE), max=max(TimeLag,na.rm = TRUE))
expected_sd128 <- median(sd128$TimeLag, na.rm=TRUE) * length(sd128$TimeLag)
paste("Median time * # timepoints =",expected_sd128, "s expected time required")
actual_sd128 <- sum(sd128$TimeLag, na.rm=TRUE)
paste("Sum of time intervals =", actual_sd128, "s actual time taken")
delay_sd128 <- (actual_sd128 - expected_sd128)/60
paste("Actual time - expected time =",delay_sd128,"min added in 2000-frame experiment")
```