---
title: "Assisted Use in the Confocal Microscopy Facility"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---

Data was downloaded from iLab for 1 year, covering March 1, 2017 - March 1, 2018, based on "purchase date."

Each row in the table represents one reservation or billed service request.
```{r message=FALSE, warning=FALSE, results='hide'}
library(dplyr)
library(readr)

datapath <- "/Users/confocal/github_theresaswayne/data-analysis/R/usage_data"
datafile <- "03012017_03012018_charges_report_source_data.csv"
usage <- read_csv(file.path(datapath, datafile))
```

# Collecting data by usage type

For each reservation, the level of assistance is recorded in the "Usage Type" column.

The possible values in this column are shown below.

```{r results='markup'}
usage$`Usage Type` <- factor(usage$`Usage Type`)
levels(usage$`Usage Type`)
```

Note that Usage Type "1" indicates the Image Processing service.

To get all the hours when a staff member was working with, or for, a user, we need to include *Assisted Use*, *Training*, and *1*.

1. First we **filter** the rows that contain these patterns.
1. Then we **select** the column corresponding to the hours of use.
1. Finally we **sum** all the hours.

```{r}
hrs_assist <- dplyr::filter(usage, grepl("Assisted|Training|1", `Usage Type`)) %>%
                dplyr::select(Quantity) %>%
                sum()
```

Similarly, to get hours when staff was *not* directly assisting, we select for all rows that contain the words *Unassisted* or *Independent*.
```{r}
hrs_unass <- dplyr::filter(usage, grepl("Unassisted|Independent", `Usage Type`)) %>%
                dplyr::select(Quantity) %>%
                sum()
```

# Summary

The hours and percentage of the total are shown below.
```{r}
usage_type <- c("Assisted/Service", "Non-Assisted")
hours <- c(hrs_assist, hrs_unass)

hrs_summ <- cbind(`Usage Type` = usage_type, Hours = hours) %>%
  as.data.frame(stringsAsFactors = FALSE)

hrs_summ$Hours <- as.numeric(as.character(hrs_summ$Hours)) # force R to see the hours as numeric

hrs_summ <- mutate(hrs_summ, 
                Percent = round((Hours / sum(Hours)), 2))

hrs_summ
                
```

# Predicted changes in income

If we raise only the assisted rates by a certain amount,
and assuming usage remains constant, 
we can project our increase in income.

```{r}
assisted_rate_incr <- c(0,10,20,30,40)
annual_revenue_added <- hrs_assist * assisted_rate_incr
revenue_summ <- cbind(`Hourly Rate Increase` = assisted_rate_incr, `Yearly Revenue Increase` = annual_revenue_added) %>%
  as.data.frame(stringsAsFactors = FALSE)
revenue_summ
```